Minimal Positive Influence of Helpline Efforts on Reduction of Homeless Death in Toronto

### 2024-03-12 paper/paper.qmd
fixed a few typos, still can't render. unknown error

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Sandy Yu
thanks: "Code and data are available at: https://github.com/Jingying-yu/help-line-homeless-death-analysis.git"
date: today
date-format: long
abstract: "This paper analyzes the influence of Central Intake Line staff efforts on the reduction of homeless death count in Toronto. Data on coded number of calls from Central Intake Call Wrap-Up Codes dataset and data on monthly homeless death count from Deaths of People Experiencing Homelessness dataset are used in generalized linear regression models to estimate the influence. We find that the staffs’ net efforts create a statistically insignificant influence on the reduction of homeless death, with specific types of efforts appearing counterproductive. This implies that while a positive influence does exist, from a statistical point of view, the impact of the Central Intake Line staff is not significantly different from non-existent."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

@@

library(tidyverse)
library(rstanarm)
library(arrow)
library(knitr)
library(ggplot2)
```

@@

All data used in this paper are obtained through OpenDataToronto Portal [@OpenDataToronto]. Two different datasets: Central Intake Call Wrap-Up Codes Data [@Central_Intake_Call] and Deaths of People Experiencing Homelessness[@Homeless_Death_Count], are retrieved to analyze the effect of Toronto’s Central Call Line (denoted as *Helpline* for the remainder of this paper) efforts to the death counts of homeless individuals in Toronto. Data is cleaned and analyzed using the open source statistical programming language R [@citeR] and supporting packages tidyverse [@tidyverse], janitor [@janitor], dplyr [@dplyr], rstanarm [@rstanarm], arrow [@arrow], ggplot2 [@ggplot2], and knitr [@knitr]. Detailed description of each dataset can be found in the subsections below.

@@

```{r}
#| label: tbl-monthly-coded
#| tbl-cap: Sample of monthly summation for Helpline Coded Calls data
#| echo: false

cleaned_data <- read_parquet(here::here("data/analysis_data/cleaned_data.parquet"))

@@

$mu_i = \alpha + \phi_i$

@@

```{r}
#| label: fig-total_coded
#| fig-cap: Visualization to show correlation between number of Calls Coded and Homeless Death Counts
#| echo: false
#| message: false

@@

@@

## First discussion point {#sec-first-point}

## Second discussion point

## Third discussion point

@@

Weaknesses and next steps should also be included.

@@

In @fig-total_coded_model_ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-total_coded_model_ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-total_coded_model_ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

@@

In @fig-referred_informed_model_ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-referred_informed_model_ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-referred_informed_model_ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

@@

@fig-total_coded_model_stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-total_coded_model_stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-total_coded_model_stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm for total_coded_model"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

@@

@fig-referred_informed_model_stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-referred_informed_model_stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-referred_informed_model_stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm for referred_informed_model"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

>>>>

---
title: "Minimal Positive Influence of Helpline Efforts on Reduction of Homeless Death in Toronto"
author: 
  - Sandy Yu
thanks: "Code and data are available at: https://github.com/Jingying-yu/help-line-homeless-death-analysis.git"
date: today
date-format: long
abstract: "This paper analyzes the influence of Central Intake Line staff efforts on the reduction of homeless death count in Toronto. Data on coded number of calls from Central Intake Call Wrap-Up Codes dataset and data on monthly homeless death count from Deaths of People Experiencing Homelessness dataset are used in generalized linear regression models to estimate the influence. We find that the staffs’ net efforts create a statistically insignificant influence on the reduction of homeless death, with specific types of efforts appearing counterproductive. This implies that while a positive influence does exist, from a statistical point of view, the impact of the Central Intake Line staff is not significantly different from non-existent."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
editor: 
  markdown: 
    wrap: 72
---

@@

library(tidyverse)
library(readr)
library(rstanarm)
library(knitr)
library(ggplot2)
```

@@

All data used in this paper are obtained through OpenDataToronto Portal
[@OpenDataToronto]. Two different datasets: Central Intake Call Wrap-Up
Codes Data [@Central_Intake_Call] and Deaths of People Experiencing
Homelessness[@Homeless_Death_Count], are retrieved to analyze the effect
of Toronto’s Central Call Line (denoted as *Helpline* for the remainder
of this paper) efforts to the death counts of homeless individuals in
Toronto. Data is cleaned and analyzed using the open source statistical
programming language R [@citeR] and supporting packages tidyverse
[@tidyverse], janitor [@janitor], rstanarm [@rstanarm], arrow [@arrow],
ggplot2 [@ggplot2], and knitr [@knitr]. Detailed description of each
dataset can be found in the subsections below.

@@

```{r}
#| label: tbl-monthly-coded
#| tbl-cap: Sample of monthly summation for Helpline Coded Calls data
#| echo: false
#| warning: false

cleaned_data <- read_csv(here::here("data/analysis_data/cleaned_data.csv"))

@@

$\mu_i = \alpha + \phi_i$

@@

-   there exist a net negative correlation between helpline efforts and
    homeless death count in Toronto

```{r}
#| label: fig-total-coded
#| fig-cap: Visualization to show correlation between number of Calls Coded and Homeless Death Counts
#| echo: false
#| message: false

@@

Hypothesis 1: - null hypothesis: true effect of total_calls_coded = 0 -
use t-stat to check at 95% significance level

Hypothesis 2: - null hypothesis: true effect of referred = 0 - use
t-stat to check at 95% significance level

Hypothesis 3: - null hypothesis: true effect of referred = 0 - use
t-stat to check at 95% significance level

@@

## Realistic Interpretation of Net Efforts

-   On average, for every increase in calls coded by the helpline, there
    is a 0.0003 unit decrease in homeless death in Toronto. There is an
    overall reduction, albeit very small, in homeless death as the
    number of total coded calls increase. This implies that the helpline
    need to take at minimum 3333 calls per month to save 1 homeless life
    in Toronto. On average, the helpline takes 14828.59 calls per month,
    meaning that 4.45 lives are saved on average.
-   4.45 is not a significant number in both a practical and statistical
    sense. To improve the return on total number of calls coded, we
    attempted to identify promising call codes that contribute to the
    net positive reduction in homeless death. Out of the 13 different
    call codes, *Referral to Shelter* and *Information on Homeless
    Prevention* are best suited as individual efforts of measurement for
    positive impacts because they provide a count of the number of
    callers provided with directions or advice.

## Individual Effort: Referral to Shelter

-   On average, for every homeless caller referred to shelter, there is
    a 0.0059 unit **increase** in homeless death in Toronto. Though not
    significant on a statistical level, this trend is still very
    counter-intuitive discouraging. In the case where the correlation is
    more significant, we would want to further analyze factor related to
    referrals to shelter. Accuracy on the current status of shelters,
    level of detail on instructed, and misjudgment of shelter
    eligibility are possible factors of examination. However, since the
    effect of this call code is statistically insignificant, it is more
    likely just a fluke caused by the dataset.

## Individual Effort: Information on Homelessness Prevention

@@

Weaknesses: very small sample size, only 32 observations Next steps:
investigate into the possible factors that caused a net zero impact on
reduction of homeless death count and propose possible solution using
data.

@@

In @fig-total-coded-model-ppcheckandposteriorvsprior-1 we implement a
posterior predictive check. This shows...

In @fig-total-coded-model-ppcheckandposteriorvsprior-2 we compare the
posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-total-coded-model-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

@@

In @fig-referred-informed-model-ppcheckandposteriorvsprior-1 we
implement a posterior predictive check. This shows...

In @fig-referred-informed-model-ppcheckandposteriorvsprior-2 we compare
the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-referred-informed-model-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

@@

@fig-total-coded-model-stanareyouokay-1 is a trace plot. It shows...
This suggests...

@fig-total-coded-model-stanareyouokay-2 is a Rhat plot. It shows... This
suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-total-coded-model-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm for total_coded_model"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

@@

@fig-referred-informed-model-stanareyouokay-1 is a trace plot. It
shows... This suggests...

@fig-referred-informed-model-stanareyouokay-2 is a Rhat plot. It
shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-referred-informed-model-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm for referred_informed_model"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2
